My first R Markdown file
===========

This is my first R markdown file.

Here, we're going load some data.

```{r}
library(datasets)
data(airquality)
summary(airquality)
```

Let's first make a pairs plot of the data.
```{r}
pairs(airquality)
```

Here's a regression model of ozone on wind, solar radiation and temperature.

```{r}
library(stats)
fit <- lm(Ozone ~ Wind + Solar.R + Temp, data = airquality)
summary(fit)
```

Here's an unordered list:

* First element

* Second element


## Hide code or result
```{r simulation, echo = FALSE}
set.seed(1)
x <- rnorm(100)
mean(x)
```

```{r simulation2, echo = FALSE, results = "hide"}
set.seed(1)
x <- rnorm(100)
mean(x)
```

## Example of inline text

```{r computetime, echo = FALSE}
time <- format(Sys.time(), "%a %b %d %X %Y")
rand <- rnorm(1)
```

The current time is `r time`, My favorite random number is `r rand`.

## Incorporating Graphics
Let's first simulate some data.
```{r simulatedata, echo = TRUE}
x <- rnorm(100); y <- x + rnorm(100, sd = 0.5)
```
Here is a scatterplot of data.
```{r scatterplot, fig.height=4}
par(mar = c(5, 4, 1, 1), las = 1)
plot(x, y, main = "My Simulated Data")
```

## Making Tables with xtable
```{r fitmodel}
library(datasets)
data(airquality)
fit <- lm(Ozone~ Wind + Temp + Solar.R, data = airquality)
```
Here is a table of regression coefficients.
```{r showtable, results="asis"}
library(xtable)
xt <- xtable(summary(fit))
print(xt, type = "html")
```

## Setting Global Options
```{r setoptions, echo = FALSE}
library(knitr)
opts_chunk$set(echo = FALSE, results = "hide")
```
First simulate data
```{r simulatedata2, echo = TRUE}
x <- rnorm(100); y <- x + rnorm(100, sd = 0.5)
```
Here is a scatterplot of the data.

```{r scatterplot2, fig.height = 4}
par(mar = c(5, 4, 1, 1), las = 1)
plot(x, y, main = "My Simulated data")
```

```{r}
num <- 50
```
The value of "num" is `r num`